---
title: "Group Project Report"
author: "Jim Leach"
date: "22 March 2016"
output: word_document
---

```{r getData, echo = FALSE, message=FALSE, warning=FALSE}
library(MASS)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(GGally)


data <- read_csv("./data/group_exercise_data.csv")
colnames(data) <- make.names(colnames(data))     

```

You can also embed plots, for example:

```{r set_up_plots, echo=FALSE}
field_plot <- function(field) {
  ggplot(data, aes_string(x = "Day", y = field)) +
    geom_line(colour = "grey") +
    geom_smooth() +
    theme(legend.position = "bottom",
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.text.x = element_text(size = 16, colour = "black"),
          legend.text = element_text(size = 16),
          legend.title = element_text(size = 16),
          title = element_text(size = 16),
          strip.text = element_text(size = 16, colour = "black"),
          strip.background = element_rect(fill = "white"),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
          panel.grid.minor.y = element_line(colour = "lightgrey", linetype = "dotted"),
          panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
          panel.margin.y = unit(0.1, units = "in"),
          panel.background = element_rect(fill = "white", colour = "lightgrey"),
          panel.border = element_rect(colour = "black", fill = NA))
}
```

# Plots of variables by day

```{r tests, echo = FALSE}
field_plot("Number.of.Blood.Tests") + ylab("Blood tests")
```
```{r women, echo = FALSE}
field_plot("X..Women") + ylab("Women tested (%)")
```

```{r kids, echo = FALSE}
field_plot("X..Children") + ylab("Children tested (%)")
```

```{r ins, echo = FALSE}
field_plot("X..Covered.by.insurance") + ylab("Insurance covered (%)")
```

# Pairs plot

```{r pairs, echo = FALSE, fig.height = 9, fig.width=9}
ggpairs(data)
```

# Regression write-up

As the variance of the outcome (number of blood tests) was significantly higher than the mean (see table), we chose a negative binomial model over a Poisson (as negative binomial is more appropriate in this setting).

```{r var_means, echo = FALSE}
vars <- apply(data, 2, var)
means <- apply(data, 2, mean)
df <- data_frame(feature = colnames(data),
                 variance = vars,
                 means = means)
library(knitr)
kable(df, col.names = c("Feature", "Variance", "Mean"),
      caption = "Variance and means for data set features")
```

We used stepwise regression to find the best model from all features in the data. Having done so we obtained an error measure from the model using mean absolute deviation.

```{r nb, echo = F, message = FALSE, warning = FALSE}
fit <- glm.nb(Number.of.Blood.Tests ~ ., data = data)
fit <- step(fit, trace = 0)
predicted <- predict(fit, type = "response")

# Estimate model error
mad <- mean(abs(data$Number.of.Blood.Tests - predicted))
mse <- mean((data$Number.of.Blood.Tests - predicted)^2)
```

We found that, for a negative binomial model, the mean absolute deviation was `r round(mad, 2)`.

```{r predict, echo = FALSE}
# Create new data ---------------------------------------------------------
means <- colMeans(data)
new <- data_frame(Day = c(1000, 1010),
                  Number.of.Blood.Tests = rep(means[2], 2),
                  X..Women = rep(means[3], 2),
                  X..Children = rep(means[4], 2),
                  X..Covered.by.insurance = rep(means[5], 2))

# Predict new points ------------------------------------------------------
predictions <- predict(fit, newdata = new, type = "response")
```

We used this model to make predictions for the new days. As the other features were unobserved on these days, we used the mean value of each feature as a replacement. This produced the following predictions.

```{r show_predictions, echo = FALSE}
predictions %>% as.data.frame() %>% setNames("Prediction") %>% 
  cbind(data.frame(Day = c(1000, 1010))) %>% 
  dplyr::select(Day, Prediction) %>% 
  kable()
```
