---
title: "Exercise 2"
author: "Jim Leach"
date: "22 March 2016"
output: html_document
---

```{r message = FALSE}
library(quanteda)
library(dplyr)
```

# Getting used to dictionaries

## Create a dictionary

Creating a dictionary is easy with `quanteda`.

```{r}
posDict <- dictionary(list(articles = c("the", "a", "an"),
                           conjunctions = c("and", "but", "or", "nor", "for", "yet", "so")))
```

We can then apply this dictionary to a corpus of text to extract the numbers of articles and conjunctions in each text in that corpus.

```{r}
posDfm <- dfm(inaugCorpus, dictionary = posDict)
```

We can weight this by relative frequency to understand a little more:

```{r}
dfm_weight <- weight(posDfm, type = "relFreq")
```

We can also plot the values of articles:

```{r}
library(ggplot2)
dfm_weight %>% as.data.frame() %>% add_rownames() %>% 
  rename(text = rowname) %>% 
  mutate(year = substr(text, 1, 4)) %>% 
  ggplot(aes(x = as.numeric(year), y = articles)) +
  # geom_bar(stat = "identity") +
  geom_point() +
  geom_smooth()
```

And conjunctions

```{r}
dfm_weight %>% as.data.frame() %>% add_rownames() %>% 
  rename(text = rowname) %>% 
  mutate(year = substr(text, 1, 4)) %>% 
  ggplot(aes(x = as.numeric(year), y = conjunctions)) +
  geom_point() +
  geom_smooth()
```

The distribution of normalised articles is relatively normal across the years, but conjunctions have gradually increased over time.

## Heirarchical dictionaries

We can also create heirarchical dictionaries easily enough.

```{r}
heir_dict <- list(articles = list(definite = "the",
                                  indefinite = c("a", "an")),
                  conjunctions = list(coordinating = c("and", "but", "or", "nor", "for", "yet", "so"),
                                      subordinating = c("although", "because", "since", "unless")))
```

```{r}
df_heir <- dfm(inaugCorpus, dictionary = heir_dict)
```

We can then compare this approach with the previous method.

```{r}
articles_heir <- selectFeatures(df_heir, "*articles*") %>% rowSums() %>% sum()
articles <- selectFeatures(posDfm, "*articles*") %>% rowSums() %>% sum()

conj_heir <- selectFeatures(df_heir, "*conj*") %>% rowSums() %>% sum()
conj <- selectFeatures(posDfm, "*conj*") %>% rowSums() %>% sum()
```

The number of articles found in the corpus from the flat dictionary was `r articles`. The number from the heirarchical dictionary was `r articles_heir`.

For conjunctions the numbers were `r conj` and `r conj_heir` respectively.