---
title: "Retail Analytics Individual Assignment Report"
author: "Jim Leach"
date: "7 April 2016"
output:
  pdf_document:
    fig_caption: yes
  word_document: default
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(tidyr)
library(magrittr)
library(mlogit)
library(ggplot2)
library(GGally)
library(broom)
library(knitr)

# Set up theme object
theme_jim <-  theme(legend.position = "bottom",
                    axis.text.y = element_text(size = 16, colour = "black"),
                    axis.text.x = element_text(size = 16, colour = "black"),
                    legend.text = element_text(size = 16),
                    legend.title = element_text(size = 16),
                    title = element_text(size = 16),
                    strip.text = element_text(size = 16, colour = "black"),
                    strip.background = element_rect(fill = "white"),
                    panel.grid.minor.x = element_blank(),
                    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
                    panel.grid.minor.y = element_line(colour = "lightgrey", linetype = "dotted"),
                    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
                    panel.margin.y = unit(0.1, units = "in"),
                    panel.background = element_rect(fill = "white", colour = "lightgrey"),
                    panel.border = element_rect(colour = "black", fill = NA))
```

```{r source, echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE}
# Load packages
library(dplyr)
library(tidyr)
library(magrittr)
library(mlogit)
library(ggplot2)
library(GGally)
library(broom)
library(knitr)

# Set up theme object
theme_jim <-  theme(legend.position = "bottom",
                    axis.text.y = element_text(size = 16, colour = "black"),
                    axis.text.x = element_text(size = 16, colour = "black"),
                    legend.text = element_text(size = 16),
                    legend.title = element_text(size = 16),
                    title = element_text(size = 16),
                    strip.text = element_text(size = 16, colour = "black"),
                    strip.background = element_rect(fill = "white"),
                    panel.grid.minor.x = element_blank(),
                    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
                    panel.grid.minor.y = element_line(colour = "lightgrey", linetype = "dotted"),
                    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
                    panel.margin.y = unit(0.1, units = "in"),
                    panel.background = element_rect(fill = "white", colour = "lightgrey"),
                    panel.border = element_rect(colour = "black", fill = NA))

# source extra functions
source("./r/functions/toproper.R")
source("./r/functions/ggally_cor.R")
source("./r/functions/tidy_mlogit.R")

# Step 0 - get and clean --------------------------------------------------
source("./r/files/000_clean_and_filter.R")
source("./r/files/001_clean_brands.R")
source("./r/files/002_create_id.R")
source("./r/files/003_light_vs_heavy.R")
source("./r/files/009_coffee_wide.R")
source("./r/files/010_coffee_long.R")

# Step 1 - elasticity modelling ----------------------------------------------
source("./r/files/100_elasticity_models.R")
source("./r/files/101_clout_and_vuln_stats.R")


# Step 2 - switching matrices ---------------------------------------------
source("./r/files/201_cooccurence_matrices.R")
source("./r/files/202_switching_matrices.R")
source("./r/files/203_market_share_correlation.R")


# Step 3 - user type level modelling --------------------------------------
# shop choice
source("./r/files/301_choice_models.R")
# shop traffic
source("./r/files/302_traffic_models.R")

# Step 5 - Prepare further visualisations ---------------------------------

source("./r/files/501_vis_weekly_share_to_prop_promo.R")
source("./r/files/502_vis_clout_and_vuln_map.R")
source("./r/files/503_vis_repeat_purchase_rate.R")
source("./r/files/504_vis_weekly_price_to_share.R")
source("./r/files/505_vis_sales_by_week.R")
source("./r/files/506_vis_share_by_week.R")

```

# Description of Research

Using _[Katar Worldpanel](http://www.kantarworldpanel.com/global)_ data on one year of coffee sales, the task of this assignment was to respond to the question: 

> "How do stores compete? If one store sees an improvement in traffic and sales, how was that store able to achieve such improvements in performance? Is it from stealing from other stores, and which competitors are likely the most affected? How would you model this data considering these findings and what suggestion would you give in terms of growing future sales?"

Intermediate and final code/results are packaged with this report, and can be found online here: [https://github.com/Jim89/icl/tree/master/retail/project](https://github.com/Jim89/icl/tree/master/retail/project).

This report has been written from the perspective of Morrisons. 

## Executive summary

Morrisons competes in a number of ways. It improved its traffic through promotional sales. It had success growing sales by growing the total market _and_ by taking market share from other shops. Consumers are not found to be highly price sensitive at a shop level, but Morrisons is at risk in terms of price, having very little clout but high vulnerability.

***

# Analysis

## Approach

### Data preparation

Minor retailers were removed, and Aldi and Lidl were combined as "Discounters". Brands cleaned and grouped. Minor brands were labelled as "Other" and sub-brands of major labels were combined (e.g. all Nescafe sub-brands were labelled as "Nescafe"). 

Transaction and visit IDs were created. The former indicated individual purchases (the unique combination of date, customer, shop, brand). The latter indicated a visit (the unique combination of date, customer and shop). 

Consumers were classified as light, medium or heavy. This classification was performed using the average volume of coffee purchased per visit (as described in Leach et al, 2016). 

Data quality and validation results from earlier work (Leach et al, 2016) were also used. No further quality or validity checks were performed. 

### Market Structure

Numerical summaries and visualisations were created to understand the market structure. Total and per-shop weekly sales were charted, along with weekly market shares. The relationship between price and market share was plotted, as was the relationship between promotions and market share. 

The correlation between market share and _total sales_ was investigated to understand whether changes in market share were due to customers shopping elsewhere, or new sales in the market. Correlations between shops' market shares were investigated to understand which shops might be stealing customers from others. 

Shop-switching matrices were created to further understand which shops were competing. These results generated plots to show the proportion of repeat purchases for each shop, helping to assess which shops had more loyal customers. Shop co-occurrence matrices were also made.

### Price Elasticity

Own and cross-price elasticities were calculated using data aggregated to a daily level. Model selection was performed using stepwise regression and cross-validation, utilising mean-squared error to compare models (as described in Leach et al, 2016).  

The calculations were performed across all three customer segments as calculations at segment level lead to insignificant and strange results. Clout and vulnerability statistics were created. 

### Shop Choice

A simple (multinomial logit) model was created accounting for price, promotion and loyalty on shop choice. Loyalties were calculated based on the first 20 weeks of data. Customers who did not appear in the first 20 weeks of data were assumed to have the average loyalties of the customers in their segment. Unobserved heterogeneity (e.g. income, family size etc) was accounted for by applying the model to each customer segment separately. 

The model was limited as important data (such as total spend per visit, or demographics) were not available. There was no data for visits when coffee was _not_ purchased. Therefore the choice model was conditional on the purchase of coffee. This limited how robust the model was and therefore elasticities were not calculated from it at this stage.

There model also assumed that all consumers had access to each shop, and knew the price and promotional level at each alternative (which may not have been the case in reality).


### Shop Traffic

_Traffic_ (visits per day) was also modelled. This model was also specific to coffee sales. It related traffic to the brand purchased, the average price and promotional level for the shop/brand combination, as well as the minimum and maximum price (across all shops) for that brand on that day to model reference-price effects. 

The model was _not_ made conditional on a purchase: for days with no purchases it was assumed that no visits occurred. Consequently there were many days with zero visits, and so a standard _Poisson_ (not a zero-truncated) model was used. As with the choice model, elasticities were not calculated as the model was also limited by the data (specifically on visits that were made when coffee was not purchased).

***

## Results

### Market Structure

The total market size at a consumer segment level is reported in appendix one.

The plot of total and per-shop weekly transactions (figure one) shows that the market is relatively flat (the yellow line and smoother in the plot), as are the sales of individual shops.

```{r show_sales_by_week, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 9, fig.cap = "Total (in yellow, with smoother) and shop-level sales per week. Note the peaks in Morrisons' sales (the purple line) just prior to weeks 20 and 40."}
sales_by_week
```

Morrisons has two "spikes" in its weekly transactions, around weeks twenty and forty. Both of these spikes correspond with increased total market transactions, and none of the other shops show similar spikes at this time. 

The shop weekly market share plot (figure two) shows a similar pattern. Morrisons' market share (along with the other shops) is relatively constant. Morrisons captures between fifteen and twenty-five percent of the market on average. Two peaks are observed in Morrisons' share around weeks twenty and forty. In these weeks it is mainly Tesco that loses share, with Asda and Sainsburys making smaller losses.

```{r show_share_by_week, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 9, fig.cap = "Shop-level market share per week. Note the peaks in Morrisons' share (purple area) just prior to weeks 20 and 40."}
share_by_week
```

In order to understand what might be driving these changes, the relationship between both average price and promotional level with market share was investigated (figures three and four).

```{r show_price_to_share, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 10, fig.cap = "The relationship between average weekly price and weekly market share for each supermarket analysed. Coloured lines represent second order polynomial linear regressions of price on market share with the grey regions representing the 95% confidence interval for the fit."}
weekly_price_to_share + theme(axis.text.x = element_text(size = 12))
```

In terms of price, Morrisons' market share is highest at prices the middle of the range of prices that it has set. This is in contrast with, for example, Asda where higher prices clearly result in a lower market share. Morrisons has weeks with some of the highest average prices across the five shops, and in these weeks sees a sharp drop in its market share at these points. At lower prices, Morrisons' market share also drops, although what causes this is unclear. 

```{r show_price_to_promo, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 10, fig.cap = "The relationship between weekly proportion of sales on promotion and weekly market share for each supermarket. Coloured lines are second order polynomial linear regressions of proportion of promotional sales on market share with the grey regions showing the 95% confidence interval for the fit. The vertical dotted grey line is at 50% promotions, with points to to the right of that line indicating the majority of sales in that week were on promotion."}
weekly_share_to_prop_promo  + theme(axis.text.x = element_text(size = 12))
```

The relationship between the proportion of promotional sales and market share reveals some interesting insights. For Morrisons, there is a strong relationship between promotions and market share: promotions lead to increased market share. Such a strong relationship is not observed for the other shops. It is interesting to note that Morrisons' two best weeks (in terms of market share) were the weeks with the two highest levels of promotion.

Broadly speaking, a shop can grow its customers in two ways: increasing its market share, or increasing the market size. In order to investigate which of these approaches might be relevant for Morrisons, the relationship between total market size (sales across all shops) and weekly market share was plotted (figure five). The numerical correlation between share and total market size was also computed and is displayed in table one.

```{r total_sales_to_share, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 10, fig.cap = "The relationship between total sales and market share for each shop. Colour lines indicate the linear relationship between, with grey regions showing the 95% confidence interval for this fit."}
share_to_total_sales_plot + theme(axis.text.x = element_text(size = 12))
```

```{r total_sales_to_share_table, echo = FALSE, message = FALSE, warning = FALSE}
share_to_total_sales_cor %>% 
  kable(col.names = c("Shop", "Correlation", "p-Value"),
        caption = "Correlation between total market size and market share for each supermarket, along with statistical significance (p-value) of that correlation.")
```

Morrisons has a positive and highly significant relationship between total market size and market share. When the total market size increases, it is Morrisons that is gaining the most (its share has the highest correlation with total market size). This shows that Morrisons has (to some extent) grown its sales by increasing the size of the market. From the data it is not clear what caused this, though.

If Morrisons also steals customers from other shops, it is important to know which ones. The pairwise correlations between weekly market share were assessed (for all five shops) and are presented in figure six. 

```{r show_spm, echo = FALSE, warning = F, message = F, fig.height = 8.5, fig.width = 10, fig.cap = "Relationships in weekly market share between all five retailers (N.B. Discounters are Aldi and Lidl). The lower half of the plot shows the pair-wise combinations with individual linear relationships, the diagonal plots show the distribution of weekly market share for each shop, and the upper-right of the plot shows the correlation values between each combination of shop's market share. Significance codes for these values are indicated using asterisks (more means more significant)."}
share_pairs_all
```

It is clear that Morrisons increases its market share at the expense of Sainsburys and Tesco primarily, and also from Aldi and Lidl ("discounters"). Note that figure six shows an overall level view for all customer segments. The pairwise _patterns_ remain constant across the individual customer segments, however the numerical values of the correlations change slightly. As such the values for each consumer segment are presented in appendix two.

\pagebreak

The overall switching matrix is presented in table two and customer segment level matrices in appendix three. 

```{r show_switch_all, echo = FALSE}
rownames(switch_mat_all) <- toproper(rownames(switch_mat_all))
colnames(switch_mat_all) <- toproper(colnames(switch_mat_all))
switch_mat_all %>% kable(caption = "Switching matrix for all customers")
```

These matrices show which shops customers are switching between and hence can be used to estimate competition between shops. For the overall matrix it can be seen that customers that shopped at Morrisons also shopped at Asda and Tesco, with some switching to Sainsburys as well. It is also possible to estimate just how much switching is occurring in the market place. At the overall level, around `r scales::percent(switch_all)` of visits were "switched" visits, i.e. those in which the customer was shopping at a different shop than on their previous visit. This shows that there is some consumer movement between shops in the market. It should be noted that this value is not constant across the three consumer segments, with light, medium and heavy consumers showing `r scales::percent(switch_l)`, `r scales::percent(switch_m)`, and `r scales::percent(switch_h)` of "switched" visits, respectively (based on switching matrices in appendix three). 

\pagebreak

The repeat purchase rate for each shop for each consumer segment is shown in figure seven. Morrisons has a relatively constant repeat purchase rate across the three consumer segments (around 60%), although it is slightly lower for the heavy segment. 

```{r show_rep_purchase, echo = FALSE, message = F, warning = F, fig.height = 6, fig.width=10, fig.cap="Repeat purchase rate (as percentage of total purchases) for each shop in each consumer segment. Note that Morrisons' rate stays relatively constant across the threee consumer segments."}
repeat_purchase_rate_all
```

The co-occurrence matrices present little additional information in this setting, but could be used to complement the switching matrices for other goods, or for data covering all goods purchased.

\pagebreak

### Price Elasticitiy

Shop-level price elasticities were computed and the resulting elasticity matrix is presented in table three. The clout and vulnerability map for these elasticities is shown in figure seven.

```{r elast_mat, echo = F, message=F, warning=F}
elasticities <- elasticities %>% 
        as.data.frame() %>% 
        rename(Sainsburys = sains,
               Asda = asda,
               Tesco = tesco, 
               Discounters = aldi, 
               Morrisons = morrisons) %>% 
        as.matrix()
rownames(elasticities) <- colnames(elasticities)
elasticities %>% kable(caption = "Price elasticities for each shop in the data. Note that 'Discounters' refers to Aldi & Lidl.")
```

```{r cv_map, echo=F, message=FALSE, warning=FALSE, fig.height = 4, fig.width=10, fig.cap = "Clout and Vulnerability map for the five shops in the data."}
clout_and_vuln_map + theme(legend.position = "right",
                           legend.text = element_text(size = 16),
                           legend.title = element_text(size = 16))
```

At a shop level, consumers are price _inelastic_ (the absolute values of price elasticity are less than one), indicating that they are not particularly sensitive to price. This is in contrast with earlier work (Leach et al, 2016) showing that at _brand_ level consumers are much more price sensitive. 

Within this setting, however, it appears that Morrisons is most at risk from a price perspective. It has zero clout, meaning that its price has no effect on sales at other shops, but is highly vulnerable, meaning that the price of other shops (Tesco and Asda in this instance) can have a large effect on its sales. 

The lack of shop-level price sensitivity is likely due to limitations of the data covering purchases of coffee only. Consumers in reality are probably highly sensitive to the average cost of a trip to each shop. However, looking only at coffee sales does not show this. Overall shop price sensitivity could be better measured with richer data, covering entire basket purchases at each shop.

\pagebreak

### Shop Choice

As noted, the shop choice model is very limited by the data and therefore does not provide a rigorous estimation of which shops are preferred. 

However, it can measure the effect of price, promotions and loyalty on shop choice at a simple level. The results from each model at the segment level are presented in tables four, five, and six.

```{r choice_light, echo = FALSE}
fit <- summary(choice_light)$mfR2 %>% as.numeric() %>% round(3)
tbl <- tidy_mlogit(choice_light)
cap <- paste("Multinomial logit model coefficients and errors for light consumers. The McFadden $R^2$ value is", fit)
kable(tbl, caption = cap)
```

```{r choice_med, echo = FALSE}
fit <- summary(choice_medium)$mfR2 %>% as.numeric() %>% round(3)
tbl <- tidy_mlogit(choice_medium)
cap <- paste("Multinomial logit model coefficients and errors for medium consumers. The McFadden $R^2$ value is", fit)
kable(tbl, caption = cap)
```


```{r choice_heavy, echo = FALSE}
fit <- summary(choice_heavy)$mfR2 %>% as.numeric() %>% round(3)
tbl <- tidy_mlogit(choice_heavy)
cap <- paste("Multinomial logit model coefficients and errors for heavy consumers. The McFadden $R^2$ value is", fit)
kable(tbl, caption = cap)
```

Whilst the exact coefficient values are different in each model, the general pattern is similar. Morrisons is somewhere in the middle of consumer preference. It is preferred over Aldi and Lidl in all segments. However, for light consumers all the other shops are _more_ preferred (their coefficients are greater), whereas for medium and heavy consumers, Morrisons is preferred over Sainsburys. 

This tallies with the repeat purchase rate analysis earlier. For heavy and medium consumers Morrisons is able to compete with Asda and Tesco, but it struggles in the light consumer segment. For both light and medium consumers, increasing price is seen to have a negative impact on the likelihood of a visit. This is not the case for heavy consumers, which is unusual and is likely due to the absence of data on when purchases were _not_ made which may help to tease out the effect of price more readily.

This analysis could be expanded with further data on things such as total basket spend, availability of each product, household inventory levels, and household demographics; as well as with data on trips to the shop where a purchase of coffee was not made. 

### Shop Traffic

Shop traffic was also modelled at a segment level. There are limitations of these models as data on visits when coffee was not purchased are unavailable, and other details are missing, e.g. marketing levels, accurate price-reference effects, or the presence of loss-leader products in each shop. Were such data available, these models could be expanded and improved. For now they simply help to support the other analyses performed. 

The results from each model are presented in tables seven, eight, and nine.

```{r traffic_light, echo = FALSE}
tidy_poisson <- function(fit) {
fit %>% 
  tidy() %>% 
  mutate(term = ifelse(grepl("\\(In", term), "Intercept", term),
         term = ifelse(grepl("^shop", term), gsub("shop", "", term), term),
         term = ifelse(grepl("^brand", term), gsub("^brand", "", term), term),
         term = ifelse(grepl("avg_", term), "Avg. price", term),
         term = ifelse(grepl("promo_p", term), "Price promo.", term),
         term = ifelse(grepl("promo_u", term), "Unit promo.", term),
         term = ifelse(grepl("min", term), "Min. (reference) price", term),
         term = ifelse(grepl("max", term), "Max. (reference) price", term),
         term = toproper(term))
}  
tidy_poisson(traffic_light) %>% kable(caption = "Poisson model parameters for light consumers' traffic model.",
                                      col.names = c("Coefficient", "Estimate", "Std. Error", "t-Stat", "p-value"))  
```

\pagebreak

```{r traffic_medium, echo = FALSE}
tidy_poisson(traffic_medium) %>% kable(caption = "Poisson model parameters for medium consumers' traffic model.",
                                       col.names = c("Coefficient", "Estimate", "Std. Error", "t-Stat", "p-value"))
```

```{r traffic_heavy, echo = FALSE}
tidy_poisson(traffic_heavy) %>% kable(caption = "Poisson model parameters for heavy consumers' traffic model.",
                                       col.names = c("Coefficient", "Estimate", "Std. Error", "t-Stat", "p-value"))
```

The results are broadly similar across the three segments and tell a similar story as the choice models. For customers in the light segment, traffic is not predicted to be much higher for Morrisons than for Aldi and Lidl and the other shops are busier. For medium and heavy consumers Morrisons begins to compete with the other shops, especially Sainsburys and Asda. Price has a negative impact, indicating that visits are likely to be lower when prices are higher.

# Conclusion and Recommendations

There is a significant positive correlation between total market size and Morrisons' market share, so Morrisons may wish to exploit this by growing the market (perhaps by tempting non coffee-buying customers to purchase coffee) in order to increase their share, and thus revenue. 

Morrisons has also increased its market share at the expense of Aldi and Lidl, Sainsburys, and Tesco. Aldi and Lidl are preferred by light consumers, with whom Morrisons competes more weakly (as seen in repeat purchase rates and  the choice and traffic models). Given the reduced value of light consumers relative to medium and heavy (appendix one), Morrisons should focus on the medium and heavy segment in order to maximise their gains.

Morrisons may wish to emulate behaviours of their main competitors, Sainsburys and Tesco, in order to take further market share. This could be done through similar offers, price matching or marketing campaigns. 

The analyses in this report should be expanded and developed, especially the choice and traffic models. More relevant data  should be collected, including marketing information, trips with no purchases, or purchases of _other_ goods in addition to coffee. Such data would enhance these models and make them more robust. Elasticities could be calculated and deeper inferences made about the market structure.

***

\pagebreak

# Appendices

## Appendix One - Sales and spend by customer segment

```{r sales_by_seg, echo = FALSE}
coffee_clean %>% 
  select(house, cust_type, transaction_id, netspend) %>% 
  distinct() %>% 
  group_by(cust_type) %>% 
  summarise(sales = n(),
            spend = sum(netspend)) %>%
  arrange(-sales) %>% 
  ungroup() %>% 
  mutate(cust_type = toproper(cust_type),
         sales = scales::comma(sales),
         spend = scales::comma(spend)) %>% 
  kable(col.names = c("Customer Segment", "Total Sales", "Total Spend (£)"),
        caption = "Total sales and spend per customer segment")
```

## Appendix Two - Pairwise correlations in weekly market shares for light, medium, and heavy consumers

```{r share_correl_l, echo = FALSE}
colnames(weekly_share_cor_l) <- gsub("Discounters", "Aldi & Lidl", colnames(weekly_share_cor_l))
rownames(weekly_share_cor_l)<- colnames(weekly_share_cor_l)
weekly_share_cor_l %>% kable(caption = "Pairwise weekly market share correlations for light consumers")
```

```{r share_correl_m, echo = FALSE}
colnames(weekly_share_cor_m) <- gsub("Discounters", "Aldi & Lidl", colnames(weekly_share_cor_m))
rownames(weekly_share_cor_m)<- colnames(weekly_share_cor_m)
weekly_share_cor_m %>% kable(caption = "Pairwise weekly market share correlations for medium consumers")
```

```{r share_correl_h, echo = FALSE}
colnames(weekly_share_cor_h) <- gsub("Discounters", "Aldi & Lidl", colnames(weekly_share_cor_h))
rownames(weekly_share_cor_h)<- colnames(weekly_share_cor_h)
weekly_share_cor_h %>% kable(caption = "Pairwise weekly market share correlations for heavy consumers")
```

\pagebreak

## Appendix Three - Customer-segment level switching matrices

```{r switch_mat_l, echo = FALSE}
rownames(switch_mat_l) <- toproper(rownames(switch_mat_l))
colnames(switch_mat_l) <- toproper(colnames(switch_mat_l))
switch_mat_l %>% kable(caption = "Switching matrix for light consumers")
```

```{r switch_mat_m, echo = FALSE}
rownames(switch_mat_m) <- toproper(rownames(switch_mat_m))
colnames(switch_mat_m) <- toproper(colnames(switch_mat_m))
switch_mat_m %>% kable(caption = "Switching matrix for medium consumers")
```

```{r switch_mat_h, echo = FALSE}
rownames(switch_mat_h) <- toproper(rownames(switch_mat_h))
colnames(switch_mat_h) <- toproper(colnames(switch_mat_h))
switch_mat_h %>% kable(caption = "Switching matrix for heavy consumers")
```

## Appendix Four - Customer-segment level cooccurrence matrices

```{r cooc_mat_l, echo = FALSE}
cooccurence_l %>% kable(caption = "Cooccurrence matrix for light consumers")
```

\pagebreak

```{r cooc_mat_m, echo = FALSE}
cooccurence_m %>% kable(caption = "Cooccurrence matrix for medium consumers")
```

```{r cooc_mat_h, echo = FALSE}
cooccurence_h %>% kable(caption = "Cooccurrence matrix for heavy consumers")
```

***

# References

1. Leach et al (2016), _Retail Analytics Group Report_, Imperial College Business School, code and analysis available online [here](https://github.com/Jim89/retailGroupProject).
2. Henningsen, Arne and Toomet, Ott (2011). maxLik: A package for maximum likelihood estimation in R. Computational Statistics 26(3), 443-458. DOI 10.1007/s00180-010-0217-1.