---
title: "Statistics and Econometrics - Assignment Four"
author: "Jim Leach"
date: "5 November 2015"
output: pdf_document
---

# Document purpose

The purpose of this document is to write up the response to the fourth assignment given as part of the _Statistics and Econometrics_ module. It is divided in to two sections, each covering a section of the assignment as given. Each major section is divided in to further subsections, one for each question in the assignment.

A number of `R` packages have been used throughout this assignment. These are loaded before the responses are given:

```{r load_pkg, echo = TRUE, message = FALSE}
library(dplyr)
library(magrittr)
library(broom)
library(knitr)
library(car)
```

***

# Section 1

In this section, some educational data from the [Michigan Department of Education](www.michigan.gov/mde) are provided. Relationships between some of the key variables in this data are explored and some models created.

```{r 1load}
load("./data/meap00_01.RData")
```

