---
title: "Optimisation and Decision Models Assignement Four"
author: "Jim Leach"
date: "13 December 2015"
output: pdf_document
---

# Document Purpose

The purpose of this document is to write up the response to the fourth assignment give as part of the _Optimistaion and Decision Models_ module. It is divided in to sections, one for each question in the assignment.

### Supplementary Note

The optimisation step of this assignment was solved using a generic quadratic programming solver available in the R language (`solve.QP`). I discussed using R with Professor Weisemann and he said it was acceptable as long as I was using a generic solver and not a bespoke function for the Markovitz model. As I have used a generic solver (as opposed to a pre-built mean-variance portfolio solver) I have adhered to this restriction.

***
```{r load_pckgs, echo=FALSE, message=FALSE}
library(quantmod)
library(dplyr)
library(knitr)
```

# Markovitz Portfolio Optimisation

## Section A - Weekly price data

Price data for the following 10 stocks was downloaded from Yahoo finance using the `quantmod` package for R. 

```{r A_show_stocks, echo = FALSE}
stocks <- data.frame(co = c("Royal Dutch Shell",
                            "HSBC",
                            "BP",
                            "Vodafone Group",
                            "GlaxoSmithKline",
                            "British American Tobacco",
                            "SABMiller",
                            "Diageo",
                            "BG Group",
                            "Rio Tinto Group"),
                     ticker = c("RDSA",
                                "HSBA",
                                "BP",
                                "VOD",
                                "GSK",
                                "BTI",
                                "SAB",
                                "DGE",
                                "BG",
                                "RIO"),
                     sector = c("Oil & Gas",
                                "Banking",
                                "Oil & Gas",
                                "Telecomms",
                                "Pharmaceuticals",
                                "Tobacco",
                                rep("Beverages",2),
                                "Oil & Gas",
                                "Mining"))
kable(stocks, col.names = c("Company Name", "Ticker", "Sector"))
```

Data were extracted going back to 2013-01-01 (i.e. the last two years of data) and it was ensured that the "Adjusted Close" data were extracted in order to account for dividends and stock splits. These data are presented as tables in Appendix 1.

```{r A_get_prices, echo = FALSE, message = FALSE}
source("../code/001_get_symbols.R")
```

***

## Section B - Weekly Returns

The weekly returns were calculated from adjusted closing price using \eqref{eq:returns}.

\begin{equation} \label{eq:returns}
Return = \dfrac{Close_i - Close_{i-1}}{Close_{i-1}}
\end{equation}

Where $Close_i$ represents the adjusted (weekly) closing price in week $i$. A full set of the weekly returns data is presented in Appendix B.

The mean weekly returns were calculated for each stock and the results are presented below.

```{r A_mean_returns, echo = FALSE, message = FALSE}
get_returns <- function(fin_data){
    closes <- fin_data[, 6]
    closes <- as.numeric(closes)
    periods <- length(closes)
    returns <- sapply(2:periods, 
                      function(x) (closes[x] - closes[x-1])/closes[x-1])
    return(returns)
}  

objects <- list(RDSA.L, HSBA.L, BP.L, VOD.L, GSK.L, BATS.L, SAB.L, DGE.L, BG.L, RIO.L)
# returns <- lapply(objects, weeklyReturn)
    weekly <- lapply(objects, to.weekly)
    weekly_dfs <- lapply(weekly, data.frame)
    returns <- lapply(weekly_dfs, get_returns)
stocks <- list("RDSA.L", "HSBA.L", "BP.L", "VOD.L", "GSK.L", "BATS.L", "SAB.L", "DGE.L", "BG.L", "RIO.L") 

# combine in to single data frame with tidy names    
returns_data <- lapply(returns, data.frame) %>% 
    bind_cols %>% 
    setNames(stocks)
names(returns_data) <- make.names(names(returns_data))

# calculate mean returns  
avg_returns <- apply(returns_data, 2, mean)
avg_returns %>% 
    data.frame() %>% 
    # t() %>% 
    kable(row.names = NA, digits = 6, format.args = list(scientific = FALSE),
          col.names = "Mean Weekly Return")
```

Similarly, the covariances of the returns between all pairs of stocks were calculated, and the results are presented below.

```{r B_cov, echo = FALSE}
covars <- cov(returns_data)
covars %>% 
    data.frame() %>% 
    kable(digits = 5)
```

***

## Section C - The Efficient Frontier

(Note that all the code used to perform the analyses discussed here can be found in the attached file `efficient_frontier.R`)

The Markowitz problem can be interpretted as a specific case of the more general class of _quadratic program_ problems^1, 2^. In quadratic programming, the objective function is a convex quadratic equation that is subject to _linear_ constraints. 

The "classic" Markowitz problem for finding the minimum variance portfolio is given by \eqref{eq:mark}.

\begin{equation} \label{eq:mark}
\begin{split}
\min x^T\Sigma x \\
\text{s.t:} \\
e^T x = 1 \\
\mu^T x \geq \underline{R} \\
x \geq 0
\end{split}
\end{equation}

This equation can be reformulated to the general quadratic program \eqref{eq:quad}.

\begin{equation} \label{eq:quad}
\begin{split}
\min -\mu^T x + \frac{1}{2}x^T\Sigma x \\
\text{s.t:} \\
A^T x \geq b_0
\end{split}
\end{equation}

In these equation, $x$ is the vector of the decision variables - i.e. the proportion of each stock to purchase, $\mu$ is a vector of expected returns for each stock, $\Sigma$ is a matrix of the covariance of the returns between all stocks, $A$ is a matrix defining the constraints and $b_0$ is a vector that defines the right-hand-side of the constraints. 

Given that there are ten stocks to be chosen from: $\mu$ has dimensions $10x1$ (1 expected return for each stock) and $\Sigma$ has dimensions $10x10$ (each stocks' returns against the others, including itself). 

The constraints are such that the proportions of each stock purchased must sum to 1 (referred to as the "proportionality constraint") and additionally we cannot short sell (i.e. we cannot have a negative amount of a stock in our portfolio). Therefore the matrix $A$ has dimensions $10x11$ (1 row for each stock, 1 column that defines the proportionality constraint and 10 columns that define the no short selling constraints). The vector $b_0$ is of length 11 (1 value for the proportionality constraint and 10 values for no-short-selling on each stock)

The matrices for $\mu$ and $\Sigma$ have been defined in section B (they are the mean weekly returns and the covariances, respectively). The constraint matrix and vector, $A$ and $b_0$ are defined as:

#### A

\setcounter{MaxMatrixCols}{20}
\[
\begin{pmatrix}
1 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0  \\
1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0  \\
1 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0  \\
1 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0  \\
1 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0  \\
1 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0  \\
1 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0  \\
1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0  \\
1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0  \\
1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1  
\end{pmatrix}
\]

#### $b_0$

\[
\begin{pmatrix} 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \end{pmatrix}
\]

Performing the necessary matrix multiplication it can be seen that these matrices would enforce the constraints as listed. 

A generic quadratic problem solver in the `R` package `quadprog` can be used to solve this program. However it is also necessary to generate an efficient frontier, which requires evaluating the solution to this program at a range of $\underline{R}$ values. i.e. for a range of desired returns. In order to do so, the vector of expected returns $\mu$ can be multiplied by some factor that models this. For example, if $\mu$ is multiplied by 0 then the first term \eqref{eq:quad} dissapears, and the solver will simply return the minimum risk portfolio. Similarly, if $\mu$ is multiplied by some very large number, then the solver will return a higher-risk portfolio. 

This was performed in R using an iterative approach, solving the quadratic program a number of times in order to find the portfolio for a range of returns values. The multiplication factor of $\mu$  ranged from 0 (i.e. finding the minimum risk portfolio) to 10 (finding a portfolio with much higher returns) using increments of 0.001. In all this represents 10,001 possible portfolio combinations.

Having found the optimal solution for each of these 10,001 returns values, the return/risk efficient frontier curve was plotted for a range of portfolio compositions and is presented below. Note that Portfolio risk is measured by the variance of the portfolio, $x^T\Sigma x$ and both return and risk are weekly values.


```{r c_solve, echo = FALSE, fig.height=6, fig.width=10}
source("../code/002_calc_returns.R")
source("../code/003_plots.R")
frontier
```

From the results set, 3 portfolios were then selected (and are additionally marked on the efficient frontier graph). These portfolios are as follows:

### Low-Risk, Low-Return

Presented below is a low-risk, low-return portfolio composition with its associated variance (risk) and return. The values of each stock show the percentage volume of that stock in the portfolio.

```{r show_low, echo = FALSE, fig.height=6, fig.width=10}
cbind(round(low[, 1:10]*100, 2), low[, 11], low[, 12]) %>% 
kable(col.names = c(names(low)[1:10], "Variance (Risk)", "Returns"), 
      digits = 5)
low_plot
```

### Medium-Risk, Medium-Return

Presented below is a medium-risk, medium-return portfolio composition with its associated variance (risk) and return. The values of each stock show the percentage volume of that stock in the portfolio.

```{r show_med, echo = FALSE, fig.height=6, fig.width=10}
cbind(round(med[, 1:10]*100, 2), med[, 11], med[, 12]) %>% 
kable(col.names = c(names(low)[1:10], "Variance (Risk)", "Returns"), 
      digits = 5)
med_plot
```


### High-Risk, High-Return

Presented below is a high-risk, high-return portfolio composition with its associated variance (risk) and return. The values of each stock show the percentage volume of that stock in the portfolio.

```{r show_high, echo = FALSE, fig.height=6, fig.width=10}
cbind(round(high[, 1:10]*100, 2), high[, 11], high[, 12]) %>% 
kable(col.names = c(names(low)[1:10], "Variance (Risk)", "Returns"), 
      digits = 5)
high_plot
```

## Section D - Limited allocations per sector

If the portfolio was limited to allow only one stock to be chosen from each sector, then it would be necessary to add a further constraint to this model. This can be modelled by \eqref{eq:limit}.

\begin{equation} \label{eq:limit}
\underline{x}_J \leq \sum_{i \in J}x_i \leq \bar{x}_J
\end{equation}

# Appendices

## Appendix One - Weekly Price Data

### Royal Dutch Shell

```{r show_rds, echo = FALSE}
kable(weekly_dfs[1], col.names = c("Open", "High", "Low", "Close", "Volume", "Adjusted Close"))
```

***

### HSBC

```{r show_hsbc, echo = FALSE}
kable(weekly_dfs[2], col.names = c("Open", "High", "Low", "Close", "Volume", "Adjusted Close"))
```

***

### BP

```{r show_bp, echo = FALSE}
kable(weekly_dfs[3], col.names = c("Open", "High", "Low", "Close", "Volume", "Adjusted Close"))
```

***

### Vodafone

```{r show_vod, echo = FALSE}
kable(weekly_dfs[4], col.names = c("Open", "High", "Low", "Close", "Volume", "Adjusted Close"))
```

***

### GlaxoSmithKline

```{r show_gsk, echo = FALSE}
kable(weekly_dfs[5], col.names = c("Open", "High", "Low", "Close", "Volume", "Adjusted Close"))
```

***

### British American Tobacco

```{r show_bti, echo = FALSE}
kable(weekly_dfs[6], col.names = c("Open", "High", "Low", "Close", "Volume", "Adjusted Close"))
```

***

### SABMiller

```{r show_sab, echo = FALSE}
kable(weekly_dfs[7], col.names = c("Open", "High", "Low", "Close", "Volume", "Adjusted Close"))
```

***

### Diageo

```{r show_diag, echo = FALSE}
kable(weekly_dfs[8], col.names = c("Open", "High", "Low", "Close", "Volume", "Adjusted Close"))
```

***

### BG Group

```{r show_bg, echo = FALSE}
kable(weekly_dfs[9], col.names = c("Open", "High", "Low", "Close", "Volume", "Adjusted Close"))
```

***

### Rio Tinto

```{r show_rio, echo = FALSE}
kable(weekly_dfs[10], col.names = c("Open", "High", "Low", "Close", "Volume", "Adjusted Close"))
```

***

## Appendix B - Weekly Returns

```{r app_b_weekly_returns, echo = FALSE, message = FALSE}
kable(returns_data, digits = 4)
```

# References

1. http://www.princeton.edu/~rvdb/542/lectures/lec17.pdf - sourced 2015-12-13
2. http://www.rinfinance.com/RinFinance2009/presentations/yollin_slides.pdf  - sourced 2015-12-13
3. http://www.javaquant.net/papers/GoldfarbIdnani.pdf - sourced 2015-12-13 