---
title: "Statistics and Econometrics - Assigment Two"
author: "Jim Leach"
date: "22 October 2015"
output: pdf_document
---

# Document purpose

The purpose of this document is to write up the response to the second assigment give as part of the _Statistics and Econometrics_ module. It is divided in to two sections, each covering a section of the assigment as given. Each major section is divided in to further subsections, one for each question in the assigment.

A number of `R` packages have been used throughout this assignment. These are loaded before the responses are given:

```{r load_pkg, echo = TRUE, message = FALSE}
library(dplyr)
library(ggplot2)
```

***

# Section one - ZIP Code Pricing Disparity

The data for this section are ZIP-code level prices for a number of items at fast food restaurants. The data also constain characteristics of each ZIP-code population. The ZIP-codes are from New Jersey and Pennsylvania.  The aim is to understand if fast-food restaurants charge higher prices in areas with a higher proportion of black residents.

## Loading the data

The data are loaded from the provided `.RData` file.

```{r s1_load}
load("./data/discrim.RData")
```


