---
title: "Introduction to Business Analytics Assignment Report"
author: "Hoa Giang, Han Mao, Niccolo Valerio, Neha Sharma, Jim Leach"
date: "2 November 2015"
output: pdf_document
---

# People picking

## Problem

For this assignment the group was tasked with using network data collected from the Business Analytics course students to pick three, five-person teams for (1) design; (2) advocacy; and (3) implementation of graduation week plans.

### The catch

The picks were subject to two constraints:

* _Capacity_: Each team must have only five people, and the same individual could not be in more than one team; and
* _Chemistry_: The picks came with a budget. Each team could only use a maximum of 30 "visibility points" (referred to as VPs). These VPs were a proxy for popularity and were derived from a network set up to mimic the social structure of the students.

## Picking teams and this document

The team-picking exercise was carried out and the results presented to the students on the course. This document, therefore, presents the results of some more detailed analysis that was conducted as part of the assignment. 

Four questions were assigned that facilitated further exploration and understanding of the networks. The responses to these questions are presented in this document.

The assignment was competed using the R language. The data were read in to R and the `igraph` package (amongst others) was used to perform the network analysis.

***

\pagebreak

## Assignment Responses

```{r load_packages, message=FALSE, warning=FALSE, echo=FALSE}
  library(MASS)
  library(lsa)
  library(igraph)
  library(readxl)
  library(dplyr)
  library(magrittr)
  library(tidyr)
  library(ggplot2)
  library(knitr)
  library(broom)
  
```

```{r load_data, echo=FALSE}
# get the data
  source("./001_get_data.R")

# define key variables
  back_bone <- 1:57
  dat <- list(albert_hall_links, 
              workshop_links, 
              weekly_meeting_links, 
              urgent_meeting_links)

# create base table
  base <- people %>% 
          left_join(guest_list, by = c("id" = "rater")) %>% rename(guest_list = item) %>% 
          left_join(style, by = c("id" = "rater")) %>% rename(style = item) %>% 
          left_join(option1, by = c("id" = "rater")) %>% rename(option1 = item) %>% 
          left_join(option2, by = c("id" = "rater")) %>% rename(option2 = item)
```

### 1 - Regressions

```{r q1_define_functions, echo=FALSE}
# helper function to extract in-degree centrality
  get_indegrees <- function (data) {
    graph <- graph.data.frame(data, directed = TRUE)
    indegrees <- degree(graph, v = V(graph), mode = "in", loops = FALSE)
    return(indegrees)
  }

# convert network stat to data frame
  convert_stat_to_df <- function(stat){
    stat %>% 
      data.frame() %>%
      add_rownames() %>% 
      setNames(c("id", "stat")) %>% 
      apply(2, as.numeric) %>% data.frame %>% tbl_df
  }   
```

```{r q2_calculate_statistics, echo=FALSE}
# get properties
  indegrees   <- lapply(seq_along(dat), function(x) get_indegrees(dat[x]))

# extract in centrality in node order (1:57) for easier comparisons
  extract_in_order <- function (degrees_obj) {
   sapply(seq_along(back_bone), function(x) degrees_obj[as.character(x)]) 
  }
  ordered_degrees <- sapply(seq_along(indegrees), function(x) 
                          extract_in_order(indegrees[[x]]))

# clean up data in to tidy data frame
  ordered_degrees %<>% data.frame %>% 
                        tbl_df() %>% 
                        add_rownames() %>% 
                        setNames(c("id",
                                   "vp", 
                                   "design", 
                                   "implementation", 
                                   "advocacy"))
  ordered_degrees[is.na(ordered_degrees)] <- 0  # set missing values (i.e. no pick) to 0
```

Negative binomial regression is most suitable as the (count) data have variances greater than their means (Appendix 1). 

```{r q1_fit, echo=FALSE}
# negative binomial models
  fit_pop_to_des_nb <- glm.nb(design ~ vp, ordered_degrees)
  fit_pop_to_imp_nb <- glm.nb(implementation ~ vp, ordered_degrees)
  fit_pop_to_adv_nb <- glm.nb(advocacy ~ vp, ordered_degrees)
  
# create presentation objects
  base_caption1 <- "Regression coefficients from negative binomial regression of"
  base_caption2 <- "on visibility points (vp)"
  col_name <- c("Term", "Estimate", "Std. Error", "Statistic", "p-value")
  
  des_table <- fit_pop_to_des_nb %>% 
                tidy %>% 
                kable(caption = paste(base_caption1, "design in-degree centrality", base_caption2),
                      col.names = col_name)
  
  imp_table <- fit_pop_to_imp_nb %>% 
              tidy %>% 
              kable(caption = paste(base_caption1, "implementation in-degree centrality", base_caption2),
                    col.names = col_name)
  
  adv_table <- fit_pop_to_adv_nb %>% 
              tidy %>% 
              kable(caption = paste(base_caption1, "advocacy in-degree centrality", base_caption2),
                    col.names = col_name)  
  
```

`r des_table`

`r imp_table`

`r adv_table`

The _vp_ coefficient is the change in the log count of in-degree centrality in the other network(s) given a one-unit change in _vp_. 

All models showed a positive, statistically significant relationship between visibility points and the in-degrees centrality in the other networks. 


\pagebreak

### 2 - Cosine Similarity and flexibility

```{r q2_set_up_picks, echo=FALSE}
# set up picks data frame  
picks <- data_frame(id = rep(1:57, each = 57), pick = rep(1:57, 57)) %>% 
          left_join(albert_hall_links, by = c("id" = "rater_id", 
                                              "pick" = "rated_id")) %>% 
          rename(popularity = rating) %>% 
          left_join(workshop_links, by = c("id" = "rater_id", 
                                           "pick" = "rated_id")) %>% 
          rename(design = rating) %>%
          left_join(weekly_meeting_links, by = c("id" = "rater_id", 
                                           "pick" = "rated_id")) %>% 
          rename(implementation = rating) %>%
          left_join(urgent_meeting_links, by = c("id" = "rater_id", 
                                           "pick" = "rated_id")) %>% 
          rename(advocacy = rating)

picks[is.na(picks)] <- 0
```

```{r q2_calc_cosine, echo = FALSE}
# function to get cosine similarity for an id
  get_similarities <- function (data, node) {
      data %>% 
          filter(id == node) %>%
          select(-c(1, 2)) %>% 
          as.matrix() %>% 
          cosine
  }
  
# apply for all ids (returns list of lists)
  similarities <- lapply(seq_along(1:max(picks$id)), 
                         function(x) get_similarities(picks, x))
  
# define function to convert cosine list to matrix for use
  get_cosines <- function (cosine_obj) {
    temp <- cosine_obj %>% 
            unlist %>% 
            matrix(nrow = 4, ncol = 4)
    idx <- lower.tri(temp)
    values <- temp[idx]
    return(values)
  }
  
# get the cosine values
  cosine_values <- lapply(seq_along(similarities), 
                          function(x) get_cosines(similarities[x]))
```  

A higher similarity represents a lower flexibility.
  
The cosine similarities for each ID's picks were calculated and the mean value taken, the value was subtracted from one to give the flexibility score. 

I.e: Flexibility is simply one minus the mean similarity for all pick-vector to pick-vector comparisons for each individual.

```{r q2_aggregate, echo = FALSE}
# aggregate and simplify to give flexiblity score (median cosine value)
  flex_score <- sapply(seq_along(cosine_values), 
                       function(x) mean(cosine_values[[x]])) %>% 
                data.frame() %>%
                add_rownames() %>% 
                setNames(c("id", "flex")) %>%
                mutate(flex = 1-flex) %>% 
                tbl_df() %>% 
                mutate(z = (flex - mean(flex, na.rm = TRUE))/sd(flex, na.rm = TRUE)) %>% 
                arrange(-flex) %>% 
                apply(2, as.numeric) %>% 
                data.frame %>% tbl_df
# print pretty
  kable(flex_score, caption = "Flexibility score and Z-value for all 57 individuals in the class",
        col.names = c("ID", "Flexibility", "Z"))  
```  

\pagebreak

### 3 - Determining group leaders

```{r q3, echo = FALSE}
  teams <- read_excel("./data/analysis_normalised_20151101_ver2.xlsx", 
                      sheet = "tot") %>% 
            tbl_df %>% 
            filter(des_pick == 1 | imp_pick == 1 | adv_pick == 1) %>% 
            select(id, des_pick, imp_pick, adv_pick)

# join on to flexibility scores    
  flex_score_leaders <- flex_score %>% 
                        left_join(base, by = "id") %>% 
                        left_join(teams, by = "id") %>% 
                        mutate(additional_flex = ifelse(option1 != 0 & 
                                                        option1 != style, 1, 0)) %>% 
                        select(id, flex, z, additional_flex, option2, 
                               des_pick, imp_pick, adv_pick)
# find team leaders  
  des_lead <- flex_score_leaders %>% 
              filter(des_pick == 1) %>% 
              mutate(score_adj = flex + additional_flex + option2) %>% 
              filter(score_adj == max(score_adj, na.rm = TRUE))
  
  imp_lead <- flex_score_leaders %>% 
    filter(imp_pick == 1) %>% 
    mutate(score_adj = flex + additional_flex + option2) %>% 
    filter(score_adj == max(score_adj, na.rm = TRUE))
  
  adv_lead <- flex_score_leaders %>% 
    filter(adv_pick == 1) %>% 
    mutate(score_adj = flex + additional_flex + option2) %>% 
    filter(score_adj == max(score_adj, na.rm = TRUE))
  
leaders <-   bind_rows(des_lead, imp_lead, adv_lead) %>% 
              mutate(Team = c("Design", "Implementation", "Advocacy")) %>% 
              rename(ID = id, Flex = flex, Z = z) %>% 
              select(Team, ID, Flex, Z)
```

Qualitative information was used to enhance the flexibility score.

An individual was assigned one additional flexibility point if they chose to:

* host a party using different invitation methods;
* use a weighted mix of invitation methods;

and two points if they chose a 50/50 split mix of invitation methods.

Three IDs were selected to be the team leaders. They have the highest flexibility in each team.

```{r q3_leaders, echo = FALSE}
kable(leaders, caption = "Team leaders based on flexibility scores", 
      col.names = c("Team", "ID", "Adjusted Flexiblity Score", "Original Z"))
```

\pagebreak

### 4 ID Rankings

```{r q4_prep, echo=FALSE}
design <- read_excel("./data/design.xlsx") %>% 
            mutate(score = 0.6 * bet + 0.3 * eig + 0.1 * close,
                   ratio = score / vp) %>% 
            select(id, score, vp, ratio) %>% 
            arrange(-ratio)

# impelementation
implementation <- read_excel("./data/implementation.xlsx") %>% 
                  mutate(score = 0.1 * bet + 0.3 * eig + 0.6 * close,
                         ratio = score / vp) %>% 
                  select(id, score, vp, ratio) %>% 
                  arrange(-ratio)
  
# advocacy
advocacy <- read_excel("./data/advocacy.xlsx") %>% 
                mutate(score = 0.4 * bet + 0.4 * eig + 0.2 * close,
                       ratio = score / vp) %>%  
                select(id, score, vp, ratio) %>% 
                arrange(-ratio)
```

The "benefit" in each network were the design, implementation and advocacy scores used for picking:

* $Design = 0.6 Betweenness + 0.3 Eigenvector centrality + 0.1 Closeness$
* $Implementation = 0.1 Betweenness + 0.3 Eigenvector centrality + 0.6 Closeness$
* $Advocacy = 0.4 Betweenness + 0.4 Eigenvector centrality + 0.2 Closeness$

Having defined these formulae, the measure-of-value (which is defined as $score/visibility points$) scores are:

```{r q4_tables, echo = FALSE}
col_names <- c("ID", "Network Score", "Visibility Points", "Measure of Value (cost-benefit)")
kable(design, caption = "Design network measure of value", col.names = col_names)
kable(implementation, caption = "Implementation network measure of value", col.names = col_names)
kable(advocacy, caption = "Advocacy network measure of value", col.names = col_names)
```

***

\pagebreak

## Appendices

### Appendix One - Exploratory plot of in-degree centrality distributions

```{r appendix1, echo = FALSE}
ordered_degrees[, -1] %>% 
  rename(Popularity = vp,
         Design = design,
         Implementation = implementation,
         Advocacy = advocacy) %>% 
      gather() %>% 
      ggplot(aes(x = value)) +
      geom_histogram(aes(fill = key), colour = "white", binwidth = 1) +
      facet_grid(key ~ .) +
      xlab("In-Degree centrality") +
      ylab("Count") +
      theme(legend.position = "none",
            strip.background = element_rect(fill = "gray80"),
            panel.background = element_rect(fill = "gray99", colour = "gray75"),
            panel.grid.major.y = element_line(colour = "gray85"),
            panel.grid.major.x = element_blank())

cbind(c("Mean", "Variance"),
      rbind(apply(ordered_degrees[, 2:5], 2, mean),
      apply(ordered_degrees[, 2:5], 2, var))) %>% 
kable(caption = "Means and variances for in-degrees centrality accross the four networks",
      col.names = c("Statistic", "Popularity", "Design", "Implementation", "Advocacy"))
            
```  

As the data are count date, and the conditional variances of the distributions were much larger than the conditional means, negative binomial regression was chosen as the most suitable method for this regression problem.

\pagebreak

### Appendix Two - OLS Linear Regression Results
```{r appendix2, echo=FALSE}
# OLS linear models
  fit_pop_to_des <- lm(design ~ vp, ordered_degrees)  
  fit_pop_to_imp <- lm(implementation ~ vp, ordered_degrees)  
  fit_pop_to_adv <- lm(advocacy ~ vp, ordered_degrees)
  
# create presentation objects
  lm_base_caption1 <- "Regression coefficients from linear regression of"
  lm_base_caption2 <- "on visibility points (vp)"
  col_name <- c("Term", "Estimate", "Std. Error", "Statistic", "p-value")
  
  des_table_lm <- fit_pop_to_des %>% 
                tidy %>% 
                kable(caption = paste(lm_base_caption1, "design picks", base_caption2),
                      col.names = col_name)
  
  imp_table_lm <- fit_pop_to_imp %>% 
              tidy %>% 
              kable(caption = paste(lm_base_caption1, "implementation picks", base_caption2),
                    col.names = col_name)
  
  adv_table_lm <- fit_pop_to_adv %>% 
              tidy %>% 
              kable(caption = paste(lm_base_caption1, "advocacy picks", base_caption2),
                    col.names = col_name)   
```

`r des_table_lm`

`r imp_table_lm`

`r adv_table_lm`

Linear models all reveal a statistically significant, positive relationship between visibility points and picks elsewhere. These models were assessed with data of size `r nrow(ordered_degrees)` and have $R^2$ values of `r round(summary(fit_pop_to_des)$r.squared, 3)`, `r round(summary(fit_pop_to_imp)$r.squared, 3)`, `r round(summary(fit_pop_to_adv)$r.squared, 3)` respectively.