---
title: "Picking for Success"
author: "Hoa Giang, Han Mao, Niccolo Valerio, Neha Sharma, Jim Leach"
logo: ./logo/msc_logo.PNG
output: 
  ioslides_presentation:
    css: ./css/styles.css
    widescreen: true
---

## Intro + agenda

```{r packages, echo = FALSE, warning=FALSE, message=FALSE}
  library(readxl)
  library(dplyr)
  library(magrittr)
  library(igraph)
  library(networkD3)
```

> __"It's not _what_ you know, but _who_ you know..."__

<br>

1. Approach to picking
2. The Royal Albert Hall
3. Design
4. Implementation
5. Advocacy
6. Questions

# Approach to picking | And the Royal Albert Hall {#divider}

## Approach to picking

<br>

* Visibility points

<br>

* Network and individual properties

<br>

* Sorting and picking

## Royal Albert Hall {.flexbox .vcenter}

The _"Royal Albert Hall"_ network - a proxy for popularity. 

```{r albert_hall, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center'}
albert_hall_links     <- read_excel("./data/BA_Anonymised.xlsx", sheet = "relation 3780")
people <- read_excel("./data/BA_Anonymised.xlsx", sheet = "Attributes")
groups <- read_excel("./data/BA_Anonymised.xlsx", sheet = "Overview") %>% .[1:12, 1:2]
albert_hall_links %<>% apply(2, as.numeric) %>% data.frame() %>% tbl_df()
people %<>% apply(2, as.numeric) %>% data.frame() %>% tbl_df()
groups %<>% apply(2, as.numeric) %>% data.frame() %>% tbl_df()

nodes <- people %>% 
          left_join(groups, by = c("group" = "group_id")) %>% 
          mutate(id = as.numeric(id) -1)

# create a function for plotting the graph -------------------------------------
create_diagram <- function(links_data, nodes_data) {

# set up the graph data frame properly
  links_graph <- graph.data.frame(links_data) 
  V(links_graph)$name <- 1:57
  links <- as.data.frame(get.edgelist(links_graph))
  links$V1<-as.numeric(as.character(links$V1))
  links$V2<-as.numeric(as.character(links$V2))
  colnames(links)<-c("source","target")  
  link_list<-(links-1)
  
# make basic diagram 
  forceNetwork(Links = link_list, Nodes = nodes, Source = "source",
               Target = "target", NodeID = "id",
               Group = "group", opacity = .8, legend = FALSE, zoom = TRUE)
}

create_diagram(albert_hall_links, nodes_data = nodes)

```

# The picks | Design, Implementation, Advocacy

## Design

Two-day workshop to create the plan

<br>

* Creativity

<br>

* Betweenness

## Design picks



## Implementation

Meets weekly to refine the details

<br>

* Tighness

<br>

* Cohesion

<br>

* Closeness

## Implementation Picks

## Advocacy

Meets as needed to overcome roadblocks and resolve issues

<br>

* Influence

<br>

* Betweenness

<br>

* Exclusivity

## Advocacy picks

# Questions
