---
title: "Optimisation and Decision Models - Assigment One"
author: "Jim Leach"
date: "20 November 2015"
output: pdf_document
---

# Document purpose

The purpose of this document is to write up the response to the first assigment give as part of the _Optimistaion and Decision Models_ module. It is divided in to two sections, each covering a section of the assigment as given. Each major section is divided in to further subsections, one for each question in the assigment.

***

# Section One - Primo Insurance Company

In this section it is posited that the Primo Insurance company is introducing two new products: special risk insurance and mortgages. They expect to make £5 per sale on special risk insurance and £2 per sale on a mortage. They wish to establish the right mix of these new products in order to maximise their products within the confines of a set of work requirements, given by the below table:

```{r 3_requirements, echo = FALSE}
library(knitr)
wr <- data.frame(dept = c("Underwriting", "Admin", "Claims"),
                 sr = c(3, 0, 2),
                 m = c(2, 1, 0),
                 wh = c(2400, 800, 1200))
kable(wr, caption = "Work-hours per unit", col.names = c("Department", "Special Risk", "Mortage", "Work-hours available"))
```

## A - Formulation of a linear program for this problem

The linear program for this problem can be defined as follows:

### i. Decision variables

* __SR__ : sales quota of special risk insurance; and
* __M__ : sales quota of mortages

### ii. Objective function

\begin{equation} \label{eq:3a_of}
max \quad Profit = 5.SR + 2.M
\end{equation}

### iii. Constraints

\begin{equation} \label{eq:3a_const}
\begin{split}
Underwriting:\quad 3.SR + 2M \leq 2400 \\
Admin: \quad 0.SR + 1.M \leq 800 \\
Claims: \quad 2.SR + 0.M \leq 1200 \\
Nonnegativity: \quad SR,  M \geq 0
\end{split}
\end{equation}

## B - Graphical Method to solve the problem

The graphical method was used to solve the problem and is submitted as part of this homework in the file `hw1q3plot.jpg`.

```{r set_vals, echo = FALSE}
sr = 600
m = 300
optim = 3600
```

Using the graphical method, the optimal solution was found to be SR = `r sr`; M = `r m` which returns an objective value of `r optim`. At the optimal solution, the values of the contraints are:

* Underwriting = `r 3*sr + 2*m`, i.e. the underwriting constaint is _binding_;
* Admin = `r m`, i.e. the admin constaint is _not binding_; 
* Claims = `r 2*sr`, i.e. the claims constraint is _binding_; and
* Non-negativity: given that neither SR nor M are 0 in the optimal solution, this constraint is _not binding_.

## C Algebraic Solution to Simultaneous Equations

## D Solve the model with the Excel solver

The model was solved using the Excel solver add-in. It has been submitted as part of this homework in the file `q3_solve.xlsx`.

Using the solver, the same optimal solution as for the graphical method was found.

***

\pagebreak

# Section Two - Nature's best Frozen Foods

