---
title: "Optimisation and Decision Models - Assigment One"
author: "Jim Leach"
date: "20 November 2015"
output: pdf_document
---

# Document purpose

The purpose of this document is to write up the response to the first assigment give as part of the _Optimistaion and Decision Models_ module. It is divided in to two sections, each covering a section of the assigment as given. Each major section is divided in to further subsections, one for each question in the assigment.

***

# Section One - Primo Insurance Company

In this section it is posited that the Primo Insurance company is introducing two new products: special risk insurance and mortgages. They expect to make £5 per sale on special risk insurance and £2 per sale on a mortage. They wish to establish the right mix of these new products in order to maximise their products within the confines of a set of work requirements, given by the below table:

```{r 3_requirements, echo = FALSE}
library(knitr)
wr <- data.frame(dept = c("Underwriting", "Admin", "Claims"),
                 sr = c(3, 0, 2),
                 m = c(2, 1, 0),
                 wh = c(2400, 800, 1200))
kable(wr, caption = "Work-hours per unit", col.names = c("Department", "Special Risk", "Mortage", "Work-hours available"))
```

## A - Formulation of a linear program for this problem

The linear program for this problem can be defined as follows:

### i. Decision variables

* __SR__ : sales quota of special risk insurance; and
* __M__ : sales quota of mortages

### ii. Objective function

\begin{equation} \label{eq:3a_of}
max \quad Profit = 5.SR + 2.M
\end{equation}

### iii. Constraints

\begin{equation} \label{eq:3a_const}
\begin{split}
Underwriting:\quad 3.SR + 2M \leq 2400 \\
Admin: \quad 0.SR + 1.M \leq 800 \\
Claims: \quad 2.SR + 0.M \leq 1200 \\
Nonnegativity: \quad SR,  M \geq 0
\end{split}
\end{equation}

## B - Graphical Method to solve the problem

The graphical method was used to solve the problem and is submitted as part of this homework in the file `hw1q3plot.jpg`.

```{r set_vals, echo = FALSE}
sr = 600
m = 300
optim = 3600
```

Using the graphical method, the optimal solution was found to be SR = `r sr`; M = `r m` (the point labelled * in the attached diagram) which returns an objective value of `r optim`. At the optimal solution, the values of the contraints are:

* Underwriting = `r 3*sr + 2*m`, i.e. the underwriting constaint is _binding_;
* Admin = `r m`, i.e. the admin constaint is _not binding_; 
* Claims = `r 2*sr`, i.e. the claims constraint is _binding_; and
* Non-negativity: given that neither SR nor M are 0 in the optimal solution, this constraint is _not binding_.

## C Algebraic Solution to Simultaneous Equations

The optimal solution from part B occurs at the vertex of the feasible region that is described by the intersection of the _underwriting_ and the _claims_ constraint lines (see attached diagram). Therefore the simultanenous equations to be solved algebraically are given by \eqref{eq:q3_sim}.

\begin{equation} \label{eq:q3_sim}
\begin{split}
Underwriting: \quad 3.SR + 2.M \leq 2400 \\
Claims: \quad 2.SR \leq 1200 \\
Claims: \quad   SR \leq 600
\end{split}
\end{equation}

Solving these equations was performed:

\begin{equation} \label{eq:q3_sim_solve}
\begin{split}
3.SR + 2.M \leq 2400 \\
2SR \leq 1200 \\
\therefore SR \leq 600 \\
\\
Perform\quad subsitution\quad and\quad solve\quad for\quad M: \\
3*(600) + 2.M \leq 2400 \\
1800 + 2M \leq 2400 \\
2M \leq 600 \\
M \leq 300 \\
\\
\therefore SR = 600, \quad M = 300
\end{split}
\end{equation}

Therefore the solution found by algebraically solving the simulatenous equations \eqref{eq:q3_sim_solve} is that M = `r m` and SR = `r sr`, returning an objective value of `r optim`, i.e. identical to that found using the graphical method and that found using the Excel solver (question D).

## D Solve the model with the Excel solver

The model was solved using the Excel solver add-in. It has been submitted as part of this homework in the file `q3_solve.xlsx`.

Using the solver, the same optimal solution as for the graphical method was found.

***

\pagebreak

# Section Two - Nature's best Frozen Foods

In this section it is posited that "Natures's Best Frozen Foods" is trying to decide the right mix of four products that it produces in order to maximise its overall earnings. The four products use a similar set of ingredients, described by the table below:

```{r 4_requirements, echo = FALSE}
library(knitr)
ingred <- data.frame(ing = c("Carrots", "Mushrooms", "Green peppers",
                             "Broccoli", "Corn"),
                     sf = c(0.0625, 0.075, 0.0625, 0.05, 0.0),
                     b = c(0.05, 0.0, 0.05, 0.075, 0.075),
                     hm = c(0.0, 0.1, 0.075, 0.075, 0.0),
                     vc = c(0.0625, 0.0, 0.0625, 0.0625, 0.0625),
                     ms = c(3750, 2000, 3375, 3500, 3750))
kable(ingred, caption = "Mixes (volumes given in kg)", col.names = c("Ingredients", "Stir Fry", "Barbeque", "Heart mushrooms", "Veggie Crunch", "Montly ingredient supply"))
```

The four products, Stir Fry, Barbeque, Hearty Mushroom and Veggie Crunch contribute £0.22, £0.20, £0.18 and £0.18 to earnings, respectively. 

## A - Formulation of a linear program for this problem

### i. Decision Variables

* __SF__ : sales of stir fry product;
* __B__ : sales of barbeque product;
* __HM__ : sales of hearty mushroom product; and
* __VC__ : sales of veggie crunch product

### ii. Objective Function


\begin{equation} \label{eq:4a_of}
max \quad Earnings = 0.22SF + 0.20B + 0.18HM + 0.18VC
\end{equation}

### iii. Constraints:

\begin{equation} \label{eq:4a_const}
\begin{split}
Carrots: 0.0625SF + 0.05B + 0.0625VC \leq 3750 \\
Mushrooms: 0075SF + 0.1HM \leq 2000 \\
Green peppers: 0.0625SF + 0.05B + 0.075HM + 0.0625VC \leq 3375 \\
Broccoli: 0.05SF + 0.075B + 0.075HM + 0.0625VC \leq 3500 \\
Corn: 0.075B + 0.0625VC \leq 3750 \\
Nonnegativity: \quad SF,  B, HM, VC \geq 0
\end{split}
\end{equation}

## B - Solve the linear program using AMPL

The linear program was formulated as an AMPL model with the following form (note that the attached `q4_natures_best.mod` file also contains the full script).

```
# set up decision variables
var sf;
var b;
var hm;
var vc;

# set up function
maximize earnings: 0.22*sf + 0.2*b + 0.18*hm + 0.18*vc;

# add constraints
subject to carrots: 0.0625*sf + 0.05*b + 0.0625*vc <= 3750;
subject to mushrooms: 0.075*sf + 0.1*hm <= 2000;
subject to green_peppers: 0.0625*sf + 0.05*b + 0.075*hm + 0.0625*vc <= 3375;
subject to broccoli: 0.05*sf + 0.075*b + 0.075*hm + 0.0625*vc <= 3600;
subject to corn: 0.075*b + 0.0625*vc <= 3750;

# add non-neg constrains
subject to sf_nn: sf >= 0;
subject to b_nn: b >= 0;
subject to hm_nn: hm >= 0;
subject to vc_nn: vc >= 0;
```

Loading this model in to `AMPL` and running the `model` and `solve` commands, the optimal solution is given by the following:

```{r q4_set, echo = FALSE}
sf <-  26666.7
b <- 22333.3
hm <- 0
vc <- 9466.67
q4_optim <- 12037.33333
```

* Sales of stir fry (SF): $`r format(sf, scientific = FALSE)`$;
* Sales of barbeque (B): $`r format(b, scientific = FALSE)`$;
* Sales of hearty mushroom (HM): $`r format(hm, scientific = FALSE)`$;
* Sales of veggie crunch (VC): $`r format(vc, scientific = FALSE)`$.

The objective value at optimal solution was found to be £$`r format(round(q4_optim, 2), scientific = FALSE)`$. Running the `display` command on the constraint parameters in the model, it would appear that the Carrots and Corn contraints were both binding at the optimal solution, the other contraints were not.

## C - The value of an extra 100g green peppers

To understand what the "value" of an extra 100g of green peppers is, it is necessary to increase the availability of green peppers, recalculate the optimal solution and report the results. 

In doing so, the linear program developed earlier was relatively unchanged. The decision variables remained the same, the objective function \eqref{eq:4a_of} remained the same, only one of the constraints was changed.  The Green Peppers contraint was now \eqref{eq:q4_gp_const_c}.

\begin{equation} \label{eq:q4_gp_const_c}
Green peppers: 0.0625SF + 0.05B + 0.075HM + 0.0625VC \leq 3475
\end{equation}

Adjusting the model and re-evaluating it, also using AMPL (the attached `q4_natures_best_extra_pepper.mod` file contains the full script), returned the following results:

```{r 4c_set, echo = FALSE}
sf_4c <-  26666.7
b_4c <- 18333.3
hm_4c <- 0
vc_4c <- 14266.7
q4_optim_4c <- 12101.33333
```

* Sales of stir fry (SF): $`r format(sf_4c, scientific = FALSE)`$;
* Sales of barbeque (B): $`r format(b_4c, scientific = FALSE)`$;
* Sales of hearty mushroom (HM): $`r format(hm_4c, scientific = FALSE)`$;
* Sales of veggie crunch (VC): $`r format(vc_4c, scientific = FALSE)`$.

The new optimal value was found to be £$`r format(round(q4_optim_4c, 2), scientific = FALSE)`$, meaning that the extra 100g of green peppers had a value of approximately £$`r format(round(q4_optim_4c, 2)-round(q4_optim, 2), scientific = FALSE)`$ (i.e. the difference between the new and the original objective values at the optimal solution).