---
title: "Optimisation and Decision Models Module 1"
author: "Jim Leach"
date: "25 November 2015"
output:
  pdf_document
---

# An Introduction to Linear Programs

It is possible to define a _generic_ optimisation program:

\begin{equation} \label{eq:generic_opt}
\begin{split}
    minimise \quad f_0(x) \\
    subject\quad to: \quad \\
    \quad f_i(x) \leq 0 \quad \forall i = 1, ..., m \\
    \quad h_i(x) = 0 \quad \forall i = 1, ..., p
\end{split}
\end{equation}

In this _linear program_, all the functions, $f_i$, $h_i$ are defined as __affine__:

\begin{equation} \label{eq:aff}
f_i (x) = a_{i1}x_1 + ... + a_{in}x_n + b_i = \sum_{j=1}^n a_{ij}x_j + b_j = a_i^Tx + b_i
\end{equation}

Affine functions ([http://mathworld.wolfram.com/AffineFunction.html](http://mathworld.wolfram.com/AffineFunction.html)) are linear in their parameters and the coefficients can either be scalars or sparse matrices. The constant (here, $b$) is either scalar or a column vector.

That is, for the generic form of the optimisation problem \eqref{eq:generic_opt} all the contraints are affine and hence linear. 

Affine (linear) contraint examples:

* $5x-3y$
* $-y$
* $x+5$
* 0
* $-x+y$

Non-affine (non-linear) constraints:

* $3xy$
* $\sin{x}$
* $1/x$
* $x/y$
* $x^2$
* $2^x$

## Solving linear programs

When solving linear programs, there are some simple and obvious steps that can be followed.

An example is to use heuristics - focus on whichever single element will min/max $f_0(x)$ the most. Wwe could calculate the bottleneck constraint, i.e. the one that is most limiting and use that to min/max $f_0(x)$. This is easy to do, but not necessarily optimal. 

Therefore, when solving a linear program, we need to come up with the right _model_. This can be challening. 

### An example:

The Gemstone Tool Company (GTC) produces wrenches and pliers. Both tools are made from steel, and the production requires (i) a molding machine and (ii) an assembly machine.

GTC wants to determine the daily production of wrenches and pliers so as to maximise their contribution to earnings, subject to the daily demand limits.

```{r show_limits, echo = FALSE}
library(knitr)
limits = data.frame(limit = c("Steel required (lbs.)",
                              "Modling machine time (hrs)",
                              "Assembly machine time (hrs)",
                              "Demand limit (units)",
                              "Contribution to earnings ($/unit)"),
                    wrenches = c(1.5, 1, 0.3, 15000, 0.13),
                    pliears = c(1, 1, 0.5, 16000, 0.1),
                    avail = c(27000, 21000, 9000, NA, NA))
kable(limits, col.names = c("", "Wrenches", "Pliers", "Availability"), format.args = list(scientific = F))
```

3 steps are needed to define the model:

1. Define the decision variables;
2. Define the objective function; and
3. Define the constraints.

### Decision variables

For this problem, the decision variables are:

W = number of wrenches produced per day; and
P = number of pliers produced per day

### Objective function

For this problem, the objective function is:

\begin{equation}
maximise \quad 0.13W + 0.13P
\end{equation}

### Constraints

There are several contraints in this linear program:

* Steel: $1.5W + 1.0P \leq 27000$;
* Molding machine: $1.0W + 1.0P \leq 21000$;
* Assembly machine: $0.3W + 0.5P \leq 9000$;
* Demand limits: $W\leq 15000$, $P\leq16000$; and
* Non-negativity (cannot produce negative amounts!): $W, P\geq 0$

Solving this linear program, the _optimal solution_ is found to be $W = 12000$ and $P = 9000$ to give an objective value of $2,460

## More complex examples

Sometimes, the constraints at work in the system will be more complex than those in the simple GTC model. For example: New Bedford Steel:

NBS produces coking coal to produce steel, they have the following bids from 8 coal suppliers:

```{r show_nbs, echo = FALSE}
nbm <- data.frame(desc = c("Price ($/mt", "Union?", "Transport", "Volatility",
                           "Capacity (mt/yr)"),
                  a = c(49500, "Yes", "Rail", 15, 300),
                  b = c(50000, "Yes", "Truck", 16, 600),
                  c = c(61000, "No", "Rail", 18, 510),
                  d = c(63500, "Yes", "Truck", 20, 655),
                  e = c(66500, "No", "Truck", 21, 575),
                  f = c(71000, "Yes", "Truck", 22, 680),
                  g = c(72500, "No", "Rail", 23, 450),
                  h = c(80000, "No", "Rail", 25, 490))
library(knitr)
kable(nbm, col.names = c("", "Ashley", "Bedford", "Consol", "Dunby", "Earlam",
                         "Florence", "Gaston", "Hopt"))
```

NBS needs 1,225 mt of coking coal with an average volatility of at least 19%. At least 50% of the coal should come from union mines. At most, 650mt can be carried by rail, and at most 720mt by truck. 

How can they minimise costs? __A linear optimistaion__

### Decision variables

* A, B, C, D, E, F, G, H - the amount to purchase from Ashley, Bedford, Consol, ...

### Objective function

\begin{equation}
min \quad 49500*A + 50000*B + 61000*C + ... + 80000*H
\end{equation}

### Constraints:

* Unions: $\dfrac{A+B+D+F}{A+B+C+D+E+F+G+H} \geq 0.5$ - but this is not linear. So rearrange to give $A+B+D+F \geq 0.5*(A+B+C+D+E+F+G+H)$. It can also be rewritten as: $A + B + D + F \geq C + E + G + H$;

* Transport: Rail: $A+C+G+H \leq 650$; Trucks: $B+D+E+F\leq720$;

* Total requirements: $A+B+C+D+E+F+G+H \geq 1225$;

* Volatility: $\dfrac{0.15*A + 0.15*B + ... + 0.25*H}{A+B+C+...+H}\geq 0.19$ - again, non-linear so rearrange.

* Individual capacities: $A\leq300$, $B\leq600$, ...

* Non-negativity $A, B, C, ... \geq 0$

This problem can also be solved, it will just require more than simple heuristics or the graphical method (see later).

### A note of caution

When using variables without a non-negativity constraint, the sign of the inequality may change when re-writing equations in to linear form. 

For example $7/3 \geq 2$ can be rearranged to give $7 \geq 2*3$ but $7/-3 \geq 2$ rearranges to $7\leq 2*(-3)$ (which i snot true! Care must be taken. 

In the constraint $x/y \geq z$, if y is not restricted to non-negative values, then cross-multiplication _cannot_ be performed without changing the sign of the inequality. 

## Linear Programs in Standard Form:

A linear program can be defined in standard form \eqref{eq:sf}.

\begin{equation}\label{eq:sf}
\begin{split}
maximise \quad \sum_{j=1}^n c_j x_j \\
s.t: \quad \sum_{j=1}^n a_{ij} x_k \leq b_i \quad \quad \forall i = 1, ..., m \\
x_1, ..., x_n \geq 0
\end{split}
\end{equation}

This itself can be expressed in matrix form \eqref{eq:sf_m}

\begin{equation}\label{eq:sf_m}
\begin{split}
maximise\quad c^Tx \\
s.t: Ax \leq b \\
x \in \mathbb{R}_+^n
\end{split}
\end{equation}

In \eqref{eq:sf_m}, $c$, $x$ and $A$ are matrices of the form:


\[
C=
  \begin{bmatrix}
    c_1 \\
    . \\
    . \\
    . \\
    c_n
  \end{bmatrix}
\]


\[
X = 
\begin{bmatrix}
    x_1 \\
    . \\
    . \\
    . \\
    x_n
\end{bmatrix}
\]

\[
A = 
\begin{bmatrix}
    a_{11} & ... & a_{1n}\\
    \vdots & \ddots & \vdots \\
    a_{m1} & ... & a_{mn}
\end{bmatrix}
\]

### Constants in the objective function, or constraint LHS's

Constants in the objective function do __not__ affect the optimal $x$ and hence can be ignored during optimisation. Constants in the constraint LHS can be absorbed in the RHS:

\begin{equation}
\sum_{j=1}^n a_{ij}x_j + d \leq b_i \Leftrightarrow \sum_{j=1}^n a_{ij}x_j \leq (b_i - d)
\end{equation}

### Minimisation objective

A minimisation objective can be easily changed in to a maximisation by multiplying the objective function by -1. Once the linear program is solved, the optimal value of the objective function has to multiplied with -1 again (but not the optimal $x$).

### $\geq$ - Inequalities and Equality Constraints

A greater-than-or-equal to constraint can be replaced with a less-than-or-equal-to constraint by multiplying by -1.

\begin{equation}
\sum_{j=1}^n a_{ij}x_j \geq b_i \Leftrightarrow -\sum_{j=1}^n x_{ij}x_j \leq -b_i \Leftrightarrow \sum_{j=1}^n -a_{ij}x_j \leq -b_i
\end{equation}

Similarly, an equality constraint can be replaced with _two_ inequalities:

\begin{equation}
\sum_{j=1}^n a_{ij} x_j = b_i \Leftrightarrow 
    \begin{cases} 
        \sum_{j=1}^n a_{ij}x_j \leq b_i, \\ 
        \sum_{j=1}^n a_{ij}x_j \geq b_i 
    \end{cases}
\end{equation}

### Unrestricted Variables

Unrestrricted variables can be converted in to the difference between two non-negative variables:

\begin{equation}
\begin{split}
max \sum_{j=1, ..., n, j \neq \hat{j}} c_{ij} x_j + c_j x_{\hat{j}} \\
s.t: \sum_{j=1, ..., n, j \neq \hat{j}} a_{ij} x_j + a_{\hat{ij}}x_{\hat_j} \leq b_i \forall i \\
x_1, ..., x_{\hat{j}-1}, x_{\hat{j}+1}, ... x_n \geq 0
\end{split}
\end{equation}