---
title: "Optimisation and Decision Models Module 1"
author: "Jim Leach"
date: "8 Decembber 2015"
output:
  pdf_document
---

# Linear Optimisation Additional Examples

This section of notes covers some more esoteric/complex examples of linear programs, and how they can be applied to a range of problems.

## Example 1 - Personnel Scheduling

Imagine an airline that wants to staff its customer service agents on a shift pattern. Ideally, they would like to minimise the overall daily costs, whilst ensuring a satisfactory level of service (as dictated by minimum requirements for agents at various times of the day).

This problem can be represented by the following table:

```{r tbl1, echo = FALSE}
library(knitr)
tbl1 <- data.frame(timeperiod = c("6-8 am", "8-10 am", "10-12 am", "12-2 pm",
                                  "2-4 pm", "4-6 pm", "6-8 pm", "8-10 pm", "10-12 pm",
                                  "12-6 am", "Daily agent cost"),
                   shift1 = c(rep("x",4 ), rep("",6), 170),
                   shift2 = c("", rep("x",4 ), rep("",5), 160),
                   shift3 = c(rep("",3),rep("x",4 ), rep("",3), 175),
                   shift4 = c(rep("", 5), rep("x",4 ), "", 180),
                   shift5 = c(rep("", 8), rep("x",2), 195),
                   agents = c(48, 79, 65, 87, 64, 73, 82, 43, 52, 15, ""))
kable(tbl1, col.names = c("Time Period", "Shift 1", "Shift 2", "Shift3", "Shift 4",
                          "Shift 5", "Agents Required"))
```

Assuming a non-integer solution, this problem can summarised thus:

#### Decision variables

\begin{equation*}
S_i = \text{ Number of staff working shift, i}
\end{equation*}

#### Objective

\begin{equation*}
\min 170S_1 + 160S_2 + ... + 195S_5
\end{equation*}

#### Subject to

\begin{equation*}
\begin{split}
S_1 \geq 48 \\
S_1 + S_2 \geq 79 \\
S_1 + S_2 + S_3 \geq 65 \\
S_2 + S_3 \geq 64 \\
S_3 + S_4 \geq 73 \\
S_3 + S_4 \geq 82 \\
S_4 \geq 43 \\
S_4 + S_5 \geq 52 \\
S_5 \geq 15 \\
S_i \geq 0
\end{split}
\end{equation*}

This is easily solved in most linear programming software and returns optimal values for each shift.

***

## Example 2 - Investment Management

Imagine now that last year we purchased shares in five companies, $A, B, , C, D, E$ at prices $L_A, L_B, ..., L_E$. The _current_ prices are $T_A, T_B, ..., T_E$ and we expect next years' prices to be $N_A, ..., N_E$. We want to sell shares to generate cash $C$ whilst maximising next years value. Additionally, whenever we sell a share we pay a _transaction cost_ of 1% as well as capital gains tax of 30% on any capital gains at the time of sale. 

We can formulate this as an optimisation problem:

#### 1. Decision variables

$x_A, x_E, ..., X_E$ - the number of shares of each stock to sell.

#### 2. Objective function

$\max N_A(A - x_A) + N_B(B - x_B) + ... + N_E(E - x_E)$

(i.e. maximise next year's value by taking prices as a function of the amount we have now, less the amount we sell).

#### 3. Constraints 

__(1)__
$(T_A x_A - 0.001 T_A x_A - 0.3 x_A \max (T_A - L_A, 0)) + ... + (T_E x_E - 0.001 T_E x_E - 0.3 x_E \max (T_E - L_E, 0)) \geq C$

Note that if $T_A \leq L_A$ then capital gains is 0 and that $\max(T_A - L_A, 0)$ is still linear in the decision variables as it is simply a choice between constants.

__(2)__

$x_A, ..., x_E \geq 0$

__(3)__
$x_A \leq A, x_B \leq B, ..., x_E \leq E$

***

## Example 3 - Peformance Evaluation

In this example we will use a technique called _data envelopment analysis_. Imagine that we are managing a football team as we want to assess the performance of our players.

We could choose o measure the number of goals as a function of appearances (i.e. goals/appearances). However this is not entirely fair: more appearances leads to more fatigure (and therefore likely fewer goals). We could try to plot an efficient frontier of goals vs. appearances, however this is also unfair: defenders are expected to score fewer goals, but may have very high appearances.  

A solution is to compare players' "outputs" with their "inputs". E.g. outputs could be # goals, # assists, # shots per game, # successful passes etc and inputs could be appearances, yellow/red cards, salary etc. 

We then seek to minimise a players efficiency based on their inputs/putputs and those of the _other_ players.

E.g. imagine three players with the following profiles:

```{r show_players, echo = FALSE}
tbl2 <- data.frame(players = c("A", "B", "C"), goals = c(11, 15, 9), assists = c(2, 1, 9),
                   shots_per = c(3.3, 2.8, 2.6), passes = c(77.1, 81.9, 85.5),
                   appearances = c(15, 29, 36), yellow = c(1, 2, 5), red = c(0, 0, 0),
                   sal = c(12.2, 7.84, 7.28))
kable(tbl2, col.names = c("Player", "Goals", "Assists", "Shots/Game", "Pass Success (%)",
                          "Appearances", "Yellows", "Reds", "Salary (£m)"))
```

We then formulate a linear program for each player that minimises some measure of them:

\begin{equation*}
\begin{split}
\min E \\
\text{st:} \\
11f_A + 15f_B + 9f_C \geq 11 \\
2f_A + 1f_B + 9f_C \geq 2 \\
3.3f_A + 2.8f_B + 2.6f_C \geq 3.3 \\
77.1f_A + 81.9f_B + 85.5f_C \geq 77.1 \\
\\
15f_A + 29f_B + 36f_C \leq 15E \\
1f_A + 2f_B + 5f_C \leq 1E \\
12.1f_A + 7.84f_B + 7.28f_C \leq 12.1E \\
\\
f_i \geq 0
\end{split}
\end{equation*}

In this instance, E (termed "efficiency") is measured between 0 (inefficient) and 1 (efficient). $f_i$ measures that "fraction" of each player that we have on the team. The value of E that we obtain from each player/individual can then be used to assess their performance. Each time, the value on the RHS of the constraint is changed to that value for _that_ player/individual in order that we are saying to find a mix of players that is at least as good/bad as the player in question.

This type of analysis is no confined to sports analytics, it can also be used to evaluate the peformance of electric utlities (e.g. outputs are # customers, energy delivered, length of lines vs. inputs of man-hours, financial losses, capital bound in assets) or perhaps of police forces (e.g. outputs arrests, crimes solves vs. inputs number of crimes). 

***

## Example 4 - Regression

The classic linear regression is given by:

\begin{equation*}
y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + ... + \beta_k x_{ik} + \epsilon_i
\end{equation*}

And the least-squares minimisation by:

\begin{equation*}
\min_{\beta} ||y - X\beta||_2^2
\end{equation*}

This, however, can be very sensitive to outliers, so we may wish to use the 1-Normal, i.e:

\begin{equation*}
\min_{\beta} ||y - X\beta||_1
\end{equation*}

Note that the 1-norm is simply the aboluste value, i.e. $||[x_1, ..., x_i]||_1 = |x_1| + ... + |x_n|$.

To formulate this as a linear program we simply need to follow these steps:

### 1. Introduce auxiliary varaiables:

\begin{equation*}
\begin{split}
\min_{\beta, \theta} \theta_1 + ... + \theta_n \\
\text{st:} \\
\theta_i = |y_i - x_i^T\beta| \forall i = 1, ..., n
\end{split}
\end{equation*}

Here, $x_i^T$ denotes the i-th row of the variable matrix, $X$.

### 2. Equivalent reformulation

We relax the equality constraints

\begin{equation*}
\begin{split}
\min_{\beta, \theta} \theta_1 + ... + \theta_n \\
\text{st:} \\
\theta_i \geq |y_i - x_i^T\beta| \forall i = 1, ..., n
\end{split}
\end{equation*}

This can be done as, as we are minimising, if we pick a value of $\theta$ that is greater than $|y_i - x_i^T\beta|$ then there is a smaller value of $\theta$ that will return a _better_ objective value. As such the constraint will still resolve as an equality. 

### 3. Remove the absolute value operator

\begin{equation*}
\begin{split}
\min_{\beta, \theta} \theta_1 + ... + \theta_n \\
\text{st:} \\
\theta_i \geq y_i - x_i^T\beta \forall i = 1, ..., n \\
\theta_i \geq x_i^T\beta - y_i  \forall i = 1, ..., n
\end{split}
\end{equation*}

This is done as the absolute value can be derived from a negative or a positive value (i.e. the _actual_ $y_i$ can be above or below the prediction line).

A similar process can be followed for the infinity norm: $\min_{\beta} ||y - X\beta||_{\infty}$ where $||X||_{\infty} = \max |x_i|$